---
title: "SlimStampeRVignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SlimStampeRData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 7
  
)
```

# Introduction
This package is used to analyze SlimStampen App data. To use this package you need a .csv file with the data from the SlimStampen lessons that you want to analyze. 
```{r setup}
library(SlimStampeRData)
```


## Read data 
Give the .csv location, for example: data <- read_dataset("C:/Users/YourName/Desktop/data.csv"). When no location is given the example data set is used.
```{r}
data <- read_dataset()
```

## First look
To view the list of participants in your viewer you can use view(x$participants) instead.
```{r}
x <- dataset_stats(data)
x$participants
```


***

# Individual participants

## Reaction Time

```{r, fig.width = 8, fig.height = 8}
individual_RT(data)
```

## Rate of Forgetting

```{r, error = TRUE}
individual_ROF(data)
```

Oh no, this data set has no alpha calculated. Be aware that any columns that are called 'alpha' or 'activation' will be re-written.

```{r}
data <- calculate_alpha_and_activation(data)
```

You could also change the settings, if they were different in the experimental setup.


Now run again:

```{r, fig.width = 8, fig.height = 8}
individual_ROF(data)
```

***
# Average of participants
The average variables are plotted over the amount of repetitions a fact has been presented. These functions give an error if your data set does not have a column with the number of repetition for each trial.

```{r, error = TRUE}
average_accuracy_participants(data)
```

To solve this you can run the function... Be aware that any existing columns with the name 'repetition' will be re-written.

```{r}
data <- calculate_repetition(data)
```

Now the average functions will not return an error anymore.

## Accuracy
```{r}
average_accuracy_participants(data)
```

## Reaction Time
```{r}
average_RT_participants(data)
```

## Rate Of Forgetting
```{r}
average_ROF_participants(data)
```

***
# Average of facts

## Rate of Forgetting

```{r}
average_ROF_facts(data)
```
Explanation about legend. 
Also you can choose different labels.
```{r}
#Instead of "factText" you could also use "factAnswer"
average_ROF_facts(data, factNames = "factText")
```

***
# Possible Errors and warnings



